---
title: "p8105_hw1_jdr2191"
author: "Jess Rosenberg"
due date: "29SEP2021"
output: github_document
---

##Problem 1

First I will load the tidyverse
```{r}
library(tidyverse)
```

Problem 1
```{r}
problem1_df = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_char = c("A","B","C","D","E","F","G","H","I","J"),
  vec_factor = factor(c("purple","purple","purple","purple","blue","blue","blue",
                        "red","red","red")),
  problem1_df
)
```
Calculate Means:
```{r}
mean(pull(problem1_df, norm_samp))
mean(pull(problem1_df, norm_samp_pos))
mean(pull(problem1_df, vec_char))
mean(pull(problem1_df, vec_factor))
```
  norm_samp mean = -0.1230888
  norm_samp_pos mean = 0.4
  vec_char mean = cannot take the mean: argument is not numeric or logical: returning NA
  vec_factor mean = cannot take the mean: argument is not numeric or logical: returning NA
  
```{r results=hide}
as.numeric(pull(problem1_df, norm_samp_pos))
as.numeric(pull(problem1_df, vec_char))
as.numeric(pull(problem1_df, vec_factor))
```
The logical variable is converted to all 1s and 0s.
The chracter variable results in all NAs with a warning message.
The factor variable results in 1s, 2s and 3s for each of the levels. 

 What happens, and why? Does this help explain what happens when you try to take the mean?

##Problem 2
Load the penguins dataset:
```{r}
data("penguins", package = "palmerpenguins")

penguins
```

#Description of the penguins dataset using inline R code, including:

the data in this dataset, including names / values of important variables
the size of the dataset (using nrow and ncol)
the mean flipper length
