---
title: "p8105_hw1_jdr2191"
author: "Jess Rosenberg"
due date: "29SEP2021"
output: github_document
---

## Problem 1

First, I will load the tidyverse.
```{r load_tidyverse}
library(tidyverse)
```

Next, I will create a dataframe that includes a random sample of size 10 from a standard Normal distribution, a logical vector indicating whether elements of the sample are greater than 0, a character vector of length 10, and a factor vector of length 10, with 3 different factor “levels”. I will also ensure that this was done correctly by looking at the dataframe after I create the required variables.

```{r creating_df}
problem1_df = tibble(
  norm_samp = rnorm(10),
  norm_samp_greater0 = norm_samp > 0,
  letter_char = c("A","B","C","D","E","F","G","H","I","J"),
  color_factor = factor(c("purple","purple","purple","purple","blue","blue","blue",
                        "red","red","red")),
)
 problem1_df
```

Next, I will attempt to take the mean of each of these variables I just created. 
```{r means}
mean(pull(problem1_df, norm_samp))
mean(pull(problem1_df, norm_samp_greater0))
mean(pull(problem1_df, letter_char))
mean(pull(problem1_df, color_factor))
```

We can take the mean of both our numeric variable (norm_samp) and our logical variable (norm_samp_greater0). 

We cannot take the mean of our character variable (letter_char) or our factor variable (color_factor). When we attempt to do these, we get the following error: "the argument is not numeric or logical: returning NA."

To further investigate what is happening here, I will convert the logical, character, and factor variables to numeric variables. 
  
```{r, convert_numeric results="hide"}
as.numeric(pull(problem1_df, norm_samp_greater10))
as.numeric(pull(problem1_df, letter_char))
as.numeric(pull(problem1_df, color_factor))
```

The logical variable is converted to all 1s and 0s. This happens because our logical variable is binary (false becomes 0 and true becomes 1). When we took the mean of the logical variable, we got a result, which makes sense as we are taking the mean of 0s and 1s, which is possible.
The character variable results in all NAs with a warning message. This happens because character variables cannot easily be converted to numeric variables as the categories do not have numeric equivalents. When we tried to take the mean of then character vector we got an error, which makes sense because it is not possible to take the mean of character/text values.
The factor variable is converted into 1s, 2s and 3s, where each level corresponds to a number. We were previously unable to take the mean of the vector variable as each level was text, but we are able to convert a factor variable to a numeric variable and subsequently take its mean. 

## Problem 2
First, I will load the penguins dataset.
```{r load_penguins}
data("penguins", package = "palmerpenguins")
```

In order to describe the penguins dataset, I will run some code to find the number of observations, rows, and columns, as well as the mean value for flipper length (with flipper lengths of NA excluded).
```{r penguins_details}
view(penguins)
nrow(penguins)
ncol(penguins)
mean(pull(penguins, flipper_length_mm), na.rm = T)
```

`
The dataset includes penguin species, island, bill length and depth, flipper length, body mass, sex and year.
There are 8 columns and 344 rows.
The mean flipper length is 200.9152 mm. 
`

Next, I will generate a scatterplot of flipper_length_mm vs bill_length_mm with color points using the species variable.
```{r plot_penguins}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color=species)) + geom_point()
```

Lastly, I will save this scatterplot to my working directory.
```{r save_plot_penguins}
ggsave("penguin_plot.pdf", height = 4, width = 6)
```


