---
title: "p8105_hw1_jdr2191"
author: "Jess Rosenberg"
due date: "29SEP2021"
output: github_document
---

#Problem 1

Loading the tidyverse
```{r}
library(tidyverse)
```

Problem 1
```{r}
problem1_df = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp > 0,
  vec_char = c("A","B","C","D","E","F","G","H","I","J"),
  vec_factor = factor(c("purple","purple","purple","purple","blue","blue","blue",
                        "red","red","red")),
)
 problem1_df
```

Calculating Means
```{r}
mean(pull(problem1_df, norm_samp))
mean(pull(problem1_df, norm_samp_pos))
mean(pull(problem1_df, vec_char))
mean(pull(problem1_df, vec_factor))
```
  `norm_samp mean = -0.1230888
  norm_samp_pos mean = 0.4
  vec_char mean = cannot take the mean: argument is not numeric or logical: returning NA
  vec_factor mean = cannot take the mean: argument is not numeric or logical: returning NA`
  
```{r, results="hide"}
as.numeric(pull(problem1_df, norm_samp_pos))
as.numeric(pull(problem1_df, vec_char))
as.numeric(pull(problem1_df, vec_factor))
```
`
The logical variable is converted to all 1s and 0s. This happens because our logical variable is binary (false becomes 0 and true becomes 1). When we took the mean of the logical variable, we got a result, which makes sense as we are taking the mean of 0s and 1s, which is possible.
The chracter variable results in all NAs with a warning message. This happens because character variables cannot easily be converted to numeric variables as the categories do not have numeric equivalents. When we tried to take the mean of then character vector we got an error, which makes sense because it is not possible to take the mean of character/text values.
The factor variable is converted into 1s, 2s and 3s, where each level corresponds to a number. We were previously unable to take the mean of the vector variable as each level was text, but we are able to convert a factor variable to a numeric variable and subsequently take its mean. 
`

#Problem 2
Loading the penguins dataset:
```{r}
data("penguins", package = "palmerpenguins")
```

#Description of the penguins dataset using inline R code, including:
```{r}
view(penguins)
nrow(penguins)
ncol(penguins)
mean(pull(penguins, flipper_length_mm), na.rm = T)
```

`
The dataset includes penguin species, island, bill length and depth, flipper length, body mass, sex and year.
There are 8 columns and 344 rows.
The mean flipper length is 200.9152 mm. 
`

Making a scatterplot of flipper_length_mm vs bill_length_mm with color points using the species variable
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color=species)) + geom_point()
```

Save the scatterplot
```{r}
ggsave("penguin_plot.pdf", height = 4, width = 6)
```


